Katana: An ELF/DWARF Manipulation Tool with Hotpatching Capabilities

### 2011-05-09 doc/katana.tex
updated other forms of the documentation besides just the .org file

% Created 2011-04-18 Mon 01:44
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}

\title{Katana: An ELF/DWARF Manipulation Tool with Hotpatching Capabilities}
\author{James Oakley}
\date{18 April 2011}

@@

    A COMMAND is of the form COMMAND\_{}IDENTIFIER PARAM PARAM PARAM \ldots{}., where
    tokens are seperated by spaces and the number of PARAMs depends on
    the command.

@@

    A COMMAND\_{}IDENTIFIER is one or more words which identify a
    COMMAND. In many cases a command is identified by only one word,
    but sometimes similar commands are grouped by sharing the first
    word in their identifier.

@@

      \emph{Usage}: \texttt{dwarfscript emit [SECTION] ELF OUTFILE}\\
      \emph{Params}: SECTION must be the name (string) of the section to write as
                Dwarfscript. If not specified it defaults to
                ``.eh\_{}frame''. ELF must be an ELF object. OUTFILE must
                be a string with the name of a file to write the resulting
                Dwarfscript to.\\
      \emph{Function}: Writes the Dwarfscript representation of the given
                  SECTION from the given ELF to OUTFILE.

@@

      \emph{Usage}: \texttt{dwarfscript compile INFILE}\\
      \emph{Params}: INFILE must be a string containing the name of a file.\\
      \emph{Function}: Interprets the contents of the file named by INFILE
                  as Dwarfscript and compiles the Dwarfscript into
                  beinary form. Returns an array with 3 items
                  0: raw data for .eh\_{}frame 
                  1: raw data for .eh\_{}frame\_{}hdr
                  2: raw data for .gcc\_{}except\_{}table.
\end{itemize} % ends low level

\item replace\\
\label{sec-2.1.2.4}

@@

\item replace section\\
\label{sec-2.1.2.4.1}

      \emph{Usage}: \texttt{replace section ELF SECTION\_NAME NEW\_SECTION}
      \emph{Params}: ELF must be an ELF object. SECTION\_{}NAME must be a
                string. NEW\_{}SECTION must be either an ELF section or raw data.
      \emph{Function}: Replaces the section with the name SECTION\_{}NAME in
                  the oject ELF with the data from
                  NEW\_{}SECTION. Section headers are replaced if NEW\_{}SECTION is
                  able to provide them, but not if it is only raw data.

\item replace raw\\
\label{sec-2.1.2.4.2}

      \emph{Usage}: \texttt{replace raw ELF OFFSET NEW\_DATA}
      \emph{Params}: ELF must be an ELF object. ADDRESS must be an
                integer. NEW\_{}DATA must be raw data.
      \emph{Function}: Replaces the raw data at OFFSET in the ELF object
                  with NEW\_{}DATA. OFFSET must refer to a location in an
                  existing section.
\end{itemize} % ends low level

\item info\\
\label{sec-2.1.2.5}

@@

\item info eh\\
\label{sec-2.1.2.5.1}

@@

\item hash\\
\label{sec-2.1.2.6}

@@

\item hash elf\\
\label{sec-2.1.2.6.1}

      \emph{Usage}: \texttt{hash elf STR}
      \emph{Params}: STR must be a string.
      \emph{Function}: Prints the result of running elf\_{}hash (from libelf)
                  on the string.

@@

\item patch\\
\label{sec-2.1.2.7}

@@

\item gen\\
\label{sec-2.1.2.7.1}

@@

\item apply\\
\label{sec-2.1.2.7.2}

@@

\item ! (shell command)\\
\label{sec-2.1.2.8}

@@

\begin{itemize}
\item .text.new
     Contains new/modified functions
\item .rodata.new, .data.new
     new data
\item .unsafe\_{}functions
     Contains a simple listing (of symbol indices) of the functions in
     the binary to be patched which should not have activation records
     on the stack when patching is taking place.
\item .debug\_{}info
     Contains listings of the variables and functions which need to be
     patched using the DWARF data format. This section is standard and
     is used here with validly formatted data, but is used for
     patching instead of debugging. The use of the the .debug\_{} name is
     preserved for compatibility with libdwarf and tools such as
     readelf, objdump, dwarfdump capable of listing DWARF
     information. It can be, however, confusing and the name will
     likely change in the future.
\item .debug\_{}frame
     Like .debug\_{}info a standard section used in a nonstandard way,
     see notes above about the naming. Contains an extended version of
     DWARF Call Frame Information which describes how various data
     structures are to be patched. The details are not properly
     documented at the moment, please email the Authors for more
     details if you would like further information.
\end{itemize}
\subsection{Patch Generation Process}
\label{sec-3.7}

>>>>

% Created 2011-05-08 Sun 21:11
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\usepackage{wrapfig}
\usepackage{soul}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{graphicx}

\title{Katana: An ELF/DWARF Manipulation Tool with Hotpatching Capabilities}
\author{James Oakley}
\date{08 May 2011}

@@

    A COMMAND is of the form COMMAND$_{\mathrm{IDENTIFIER}}$ PARAM PARAM PARAM \ldots{}., where
    tokens are seperated by spaces and the number of PARAMs depends on
    the command.

@@

    A COMMAND$_{\mathrm{IDENTIFIER}}$ is one or more words which identify a
    COMMAND. In many cases a command is identified by only one word,
    but sometimes similar commands are grouped by sharing the first
    word in their identifier.

@@

      \emph{Usage}: \texttt{dwarfscript emit [SECTION] ELF OUTFILE}\\
      \emph{Params}: SECTION must be the name (string) of the section to write as
                Dwarfscript. If not specified it defaults to
                ``.eh$_{\mathrm{frame}}$''. ELF must be an ELF object. OUTFILE must
                be a string with the name of a file to write the resulting
                Dwarfscript to.\\
      \emph{Function}: Writes the Dwarfscript representation of the given
                  SECTION from the given ELF to OUTFILE.

@@

      \emph{Usage}: \texttt{dwarfscript compile INFILE}\\
      \emph{Params}: INFILE must be a string containing the name of a file.\\
      \emph{Function}: Interprets the contents of the file named by INFILE
                  as Dwarfscript and compiles the Dwarfscript into
                  beinary form. Returns an array with 3 items
                  0: raw data for .eh$_{\mathrm{frame}}$ 
                  1: raw data for .eh$_{\mathrm{frame}}$$_{\mathrm{hdr}}$
                  2: raw data for .gcc$_{\mathrm{except}}$$_{\mathrm{table}}$.
\end{itemize} % ends low level

\item extract\\
\label{sec-2.1.2.4}

@@

\item extract section\\
\label{sec-2.1.2.4.1}

      \emph{Usage}: \texttt{extract section ELF SECTION\_NAME}
      \emph{Params}: ELF must be an ELF object. SECTION$_{\mathrm{NAME}}$ must be a
                string. 
      \emph{Function}: Returns the data and header information for the
                  specified section

\item extract section$_{\mathrm{data}}$\\
\label{sec-2.1.2.4.2}

      \emph{Usage}: \texttt{extract section ELF SECTION\_NAME}
      \emph{Params}: ELF must be an ELF object. SECTION$_{\mathrm{NAME}}$ must be a
                string. 
      \emph{Function}: Like extract section except extracts only the raw
                 data stored in the section and not any header information.
\end{itemize} % ends low level

\item replace\\
\label{sec-2.1.2.5}

@@

\item replace section\\
\label{sec-2.1.2.5.1}

      \emph{Usage}: \texttt{replace section ELF SECTION\_NAME NEW\_SECTION}
      \emph{Params}: ELF must be an ELF object. SECTION$_{\mathrm{NAME}}$ must be a
      string. NEW$_{\mathrm{SECTION}}$ must be either an ELF section or raw data.
      \emph{Function}: Replaces the section with the name SECTION$_{\mathrm{NAME}}$ in
      the oject ELF with the data from NEW$_{\mathrm{SECTION}}$. Section headers
      are replaced if NEW$_{\mathrm{SECTION}}$ is able to provide them, but not if
      it is only raw data.

\item replace raw\\
\label{sec-2.1.2.5.2}

      \emph{Usage}: \texttt{replace raw ELF OFFSET NEW\_DATA}
      \emph{Params}: ELF must be an ELF object. ADDRESS must be an
                integer. NEW$_{\mathrm{DATA}}$ must be raw data.
      \emph{Function}: Replaces the raw data at OFFSET in the ELF object
                  with NEW$_{\mathrm{DATA}}$. OFFSET must refer to a location in an
                  existing section.
\end{itemize} % ends low level

\item info\\
\label{sec-2.1.2.6}

\item info eh\\
\label{sec-2.1.2.6.1}

@@

\item hash\\
\label{sec-2.1.2.7}

@@

\item hash elf\\
\label{sec-2.1.2.7.1}

      \emph{Usage}: \texttt{hash elf STR}
      \emph{Params}: STR must be a string.
      \emph{Function}: Prints the result of running elf$_{\mathrm{hash}}$ (from libelf)
                  on the string.

@@

\item patch\\
\label{sec-2.1.2.8}

@@

\item gen\\
\label{sec-2.1.2.8.1}

@@

\item apply\\
\label{sec-2.1.2.8.2}

@@

\item ! (shell command)\\
\label{sec-2.1.2.9}

@@

\begin{itemize}
\item .text.new
     Contains new/modified functions
\item .rodata.new, .data.new
     new data
\item .unsafe$_{\mathrm{functions}}$
     Contains a simple listing (of symbol indices) of the functions in
     the binary to be patched which should not have activation records
     on the stack when patching is taking place.
\item .debug$_{\mathrm{info}}$
     Contains listings of the variables and functions which need to be
     patched using the DWARF data format. This section is standard and
     is used here with validly formatted data, but is used for
     patching instead of debugging. The use of the the .debug\_{} name is
     preserved for compatibility with libdwarf and tools such as
     readelf, objdump, dwarfdump capable of listing DWARF
     information. It can be, however, confusing and the name will
     likely change in the future.
\item .debug$_{\mathrm{frame}}$
     Like .debug$_{\mathrm{info}}$ a standard section used in a nonstandard way,
     see notes above about the naming. Contains an extended version of
     DWARF Call Frame Information which describes how various data
     structures are to be patched. The details are not properly
     documented at the moment, please email the Authors for more
     details if you would like further information.
\end{itemize}
\subsection{Patch Generation Process}
\label{sec-3.7}
